\documentclass[fleqn]{article}
\usepackage{preamble}

\title{
	Differential Equations \\
	\medskip
	\large Homework 1 -- Ordinary Differential Equations
}
\author{Abraham Murciano}

\begin{document}

\maketitle

We are to solve the following linear ordinary differential equation and initial condition.
\begin{gather}
	y'(x) + 4y(x) = 2e^{3x} + 3\cos(2x) + 2x - 1 \label{ode} \\
	y(0) = 2 \nonumber
\end{gather}

We will first state the homogeneous equation.
\begin{equation*}
	y_h'(x) + 4y_h(x) = 0
\end{equation*}

Rearranging will give us the following.
\begin{gather*}
	y_h'(x) = -4 y_h(x) \\
	y_h(x) = -\frac{1}{4}y_h'(x)
\end{gather*}

Now we can find a solution to our homogeneous equation.
\begin{equation*}
	y_h(x) = A e^{-4x}
\end{equation*}

Once we have that, we can begin working on our particular solution. First we write it in the form
\begin{equation}
	y_p(x) = B e^{3x} + C \sin(2x) + D \cos(2x) + E x + F \label{particular}
\end{equation}

Now we will differentiate \(y_p(x)\).
\begin{equation}
	y_p'(x) = 3B e^{3x} + 2C \cos(2x) - 2D \sin(2x) + E \label{particular'}
\end{equation}

Now, after substituting equations \ref{particular} and \ref{particular'} into equation \ref{ode}, we can rearrange and equate the coefficients.
\begin{multline*}
	3B e^{3x} + 2C \cos(2x) - 2D \sin(2x) + E + 4(B e^{3x} + C \sin(2x) + D \cos(2x) + E x + F) \\
	= 2e^{3x} + 3\cos(2x) + 2x - 1
\end{multline*}
\begin{equation*}
	7B e^{3x} + (2C + 4D)\cos(2x) + (4C - 2D)\sin(2x) + 4E x + E + 4F = 2e^{3x} + 3\cos(2x) + 2x - 1
\end{equation*}

When comparing the coefficients of the each term, we obtain the following system of equations.
\begin{align*}
	7B = 2 \qquad & \Rightarrow \qquad B = \frac{2}{7} \\
	2C + 4D = 3 \qquad & \Rightarrow \qquad C = \frac{3}{2} - 2D \\
	4C - 2D = 0 \qquad & \Rightarrow \qquad D = \frac{3}{5}, C = \frac{3}{10} \\
	4E = 2 \qquad & \Rightarrow \qquad E = \frac{1}{2} \\
	E + 4F = -1 \qquad & \Rightarrow \qquad F = -\frac{3}{8}
\end{align*}

Thus we have a solution for \(y_p(x)\).
\begin{equation*}
	y_p(x) = \frac{2}{7} e^{3x} + \frac{3}{10} \sin(2x) + \frac{3}{5} \cos(2x) + \frac{1}{2} x - \frac{3}{8}
\end{equation*}

Finally we have a family of solutions for \(y(x)\) by adding the homogeneous and the particular solutions.
\begin{equation*}
	y(x) = A e^{-4x} + \frac{2}{7} e^{3x} + \frac{3}{10} \sin(2x) + \frac{3}{5} \cos(2x) + \frac{1}{2} x - \frac{3}{8}
\end{equation*}

All that remains is to take the initial condition into account to obtain one single solution from our family of solutions.
\begin{gather*}
	y(0) = 2 = A + \frac{2}{7} + \frac{3}{5} - \frac{3}{8} \\
	\Rightarrow A = \frac{417}{280} \approx 1.49
\end{gather*}

Therefore the final solution for \(y(x)\) is
\begin{equation*}
	y(x) = \frac{417}{280} e^{-4x} + \frac{2}{7} e^{3x} + \frac{3}{10} \sin(2x) + \frac{3}{5} \cos(2x) + \frac{1}{2} x - \frac{3}{8}
\end{equation*}

\end{document}
